\LARGE \bf TODO Name this thing

### 2012-02-16 papers/itsc_2012/root.tex
WIP fleshing out agent section of paper

% what are our goals with this paper?
% theme: easy simulation in your city, in 2 minutes!
% * demonstrate use of OSM and motivate more data collection
% * present a simulation framework
% * present a pretty naive static planning
% * motivate and demonstrate wards for hierarchial planning (but also reasoning
%    about congestion)

@@

AORTA uses microscopic simulation, modeling individual drivers as agents.
Discrete-time simulation is used, meaning an agent accelerates at a fixed rate
for the duration of a ``tick'' to achieve a new position and velocity. The tick
duration $dt$ is configurable and fixed for the simulation. Agents know this
value and use it to reason about collision avoidance and satisfying other
constraints, so setting this value to the standard ``reaction time'' of a human
driver may be appropriate. Each tick, the simulation performs the following
steps:
\begin{itemize}
  \item Introducing new agents into the map
  \item Updating the position and velocity of agents
  \item Checking for collisions
  \item Allowing each agent to react
\end{itemize}

There are two structures to detect collisions. All traversables (edges and the
straight line-approximation of turns) maintain a queue of occupying agents,
ordering them so that the agent farthest along is at the head of the queue.
A naive method of collision detection would detect an agent passing an agent on
the same queue, or entering a new queue not at the end, immediately after an
agent moves. However, agents take their steps sequentially and in an effectively
random order. Ideally they could be reordered so that agents in the front of a
queue always move first. Since the first agent may cross into the next queue,
that introduces a depdendency between traversables. A topological sort of a
directed graph containing circuits is not possible, though, and since we
guarantee connectedness, there are obviously circuits. Instead, collisions are
checked after every agent has stepped forward in time. Queues remember the
original order of agents before a step, and afterwards can verify the order is
still legal, with new agents on the end and existing agents in the same order.
Likewise, intersections check for collisions after everybody has moved by
examining the turn each agent is performing and verifying that no other agent is
simultaneously performing a conflicting turn.

invariants for agents. interface of behavior and intersection policies.

lookahead... worst-case stop dist argument is everywhere. safe spawning, too.

intersection policies... stop sign, traffic light.

@@

systems ideas.. trylocks, atomic sim update, deadlock issues

>>>>

% what are our goals with this paper?
% theme: easy simulation in your city, in 2 minutes!
% * demonstrate use of OSM and motivate more data collection
% * present a simulation framework
% * present a pretty naive static planning
% * motivate and demonstrate wards for hierarchial planning (but also reasoning
%    about congestion)
% be sure to emphasize open source!

@@

AORTA uses microscopic simulation, modeling individual drivers as agents.
Discrete-time simulation is used, meaning an agent accelerates at a fixed rate
for the duration of a ``tick'' to achieve a new position and velocity. The tick
duration $dt$ is configurable and fixed for the simulation. Agents know this
value and use it to reason about collision avoidance and satisfying other
constraints, so setting this value to the standard ``reaction time'' of a human
driver may be appropriate. Each tick, the simulation performs the following
steps:
\begin{enumerate}
  \item Introducing new agents into the map
  \item Updating the position and velocity of agents
  \item Checking for collisions
  \item Allowing each agent to react
\end{enumerate}

Agents have few enforced constraints -- they cannot exceed their physical
acceleration capabilities, travel in reverse, or travel past the end of a
traversable without appearing on the next. Intersection policies and avoiding
collisions on the same edge are the source of the other constraint in the
simulation, but it is the job of a behavior to obey. There are three main
components of the simulation, each configurable by extending default classes:
agent behaviors, routing strategies, and intersection policies. Each will be
described below.

@@

systems ideas, or limitations.. trylocks, atomic sim update, deadlock issues

limit: no lane-changing

\subsection{Spawning}

in realtime!

\subsection{Collision checking}

There are two structures to detect collisions. All traversables (edges and the
straight line-approximation of turns) maintain a queue of occupying agents,
ordering them so that the agent farthest along is at the head of the queue.
A naive method of collision detection would detect an agent passing an agent on
the same queue, or entering a new queue not at the end, immediately after an
agent moves. However, agents take their steps sequentially and in an effectively
random order. Ideally they could be reordered so that agents in the front of a
queue always move first. Since the first agent may cross into the next queue,
that introduces a depdendency between traversables. A topological sort of a
directed graph containing circuits is not possible, though, and since we
guarantee connectedness, there are obviously circuits. Instead, collisions are
checked after every agent has stepped forward in time, guaranteeing that a
time-step is atomic. Queues remember the original order of agents before a step,
and afterwards can verify the order is still legal, with new agents on the end
and existing agents in the same order.  Likewise, intersections check for
collisions after everybody has moved by examining the turn each agent is
performing and verifying that no other agent is simultaneously performing a
conflicting turn.

\section*{Agent Reactions}

Each agent has a ``behavior'' governing it. When the agent travels past the end
of an edge during its step, the behavior picks the turn to pursue. Each tick,
the behavior can make the agent react by performing one of two actions:
disappearing from the map (only when the agent is at rest and is done with its
route) or setting an acceleration for the next step. In the future, another
possible action will be initiating a lane-change or merge.

Currently, the primary behavior used by all agents is the ``route-following
behavior.'' It is a generalized, baseline behavior that guarantees not to
collide with another agent or enter an intersection at the wrong time. It avoids
some forms of deadlock (in which no agent can move due to interdependencies),
but there is currently one known flaw. This baseline behavior picks the fasrest
conservative (safe) choice of acceleration at each step. It could be easily
extended to mimic human drivers by introducing a higher reaction time delay, for
instance.

% TODO diagram for lookahead

lookahead... worst-case stop dist argument is everywhere.

\section*{Routing Strategies}

\section*{Intersection Policies}

and future plans
